Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2021-02-19 bbbq_article.tex
Moved mono- and di-peptides to supmat
Moved minor discussion point that software is tuned to predict TMHs in human to supmat
Moved noise in dbSNP to minor discussion in supmat
Improved 'Selection of SNPs'

The first query was a call to the 'gene' database for the 
term 'membrane protein' (in all fields) for the organism \textbf{Homo sapiens}.
This resulted in 1129 gene IDs.
The next query was a call to the 'gene' database 
to obtain the gene names from the gene IDs.
Per gene name, the 'dbSNP' NCBI database is queried for 
variations associated with the gene name. 
The number of variations
was limited to the first 250 variations per gene,
resulting in 282250 variations. 
A variation needs not always be a SNP,
as dbSNP also catalogs other DNA alterations, such as, among others, insertions,
deletions and tandem repeats.
We select only the variations that result in a SNP for
a single amino acid substitution.
Per SNP, the 'protein' NCBI database was queried for the
protein sequence.
Of each protein sequence, the protein topology was determined 
using PureseqTM.
From the topology and the known location of the SNP, 
we score the location (i.e. TMH or soluble protein region) 
where the change occurred.

@@

\richel{GvdB: Also irrelevant. Only 1 protein}
% \paragraph{Topology prediction of TRDD1}
%
PureseqTM does not predict the topology
of proteins that have less than three amino acids. 
The TRDD1 ('T cell receptor delta diversity 1') protein,
however, is two peptides long. 
We corrected for this by predicting that this protein
is completely cytosolic.

\richel{GvdB: Also not really relevant. This is just a minor thing, only relevant if we would see large differences between human and bacteria (but we see not)}
% \paragraph{Bacteria have a different cell membrane}
%
In this experiment we predicted epitopes that overlap with 
TMHs from a human, bacterial and viral proteome,
would these proteins be expressed in a human host.
Bacteria, however have different cell membranes and cell walls, 
hence different structural requirements for a TMH.
Both topology prediction tools were trained to recognize
human TMHs, thus we cannot be sure that
the transmembrane regions predicted in bacterial proteins
are actually part of an integral membrane protein.
For the purpose of this study, we assume the 
error in topology predictions to be unbiased way towards topology.
In other words: that a bacterial TMH is incorrectly
predicted to be absent just as often as it is incorrectly
predicted to be present elsewhere.

@@

\richel{GvdB: Also not really relevant}
% \paragraph{False positives in SNPs}
%
Regarding the evolutionary conservation of TMHs using SNPs,
again, it is estimated that approximately ten percent
of SNPs is a false positive that result from the methods to determine
a SNP. One example is that sequence variations are incorrectly
detected due to highly similar duplicated sequences \cite{musumeci2010single}.
We assume that these duplications occur as often in TMHs as in
regions around these, hence we expect this not to affect our results.

@@

\richel{GvdB: This can go to the methods or results.}
% \paragraph{Selection of SNPs}
%
We selected to analyze SNPs with the lowest SNP ID first.
This means we have a bias for picking SNPs with
an earlier discovery date.
We expect this bias to be biologically irrelevant
and we find no clear evidence of a bias, as
shown in figure \ref{fig:n_tmhs_per_protein}.
Still, the best way to interpret this figure
is open to discussion.

@@

@@

>>>>

The first query was a call to the 'gene' database for the 
term 'membrane protein' (in all fields) for the organism \textbf{Homo sapiens}.
This resulted in 1129 gene IDs.
The next query was a call to the 'gene' database 
to obtain the gene names from the gene IDs.
Per gene name, the 'dbSNP' NCBI database is queried for 
variations associated with the gene name.

The number of variations
was limited to the first 250 variations per gene,
resulting in 282250 variations. 
A variation needs not always be a SNP,
as dbSNP also catalogs other DNA alterations, such as, among others, insertions,
deletions and tandem repeats.
We select only the variations that result in a SNP for
a single amino acid substitution.

% \paragraph{Selection of SNPs}
%
We selected to analyze SNPs with the lowest SNP ID first.
This means we have a bias for picking SNPs with
an earlier discovery date.
We expect this bias to be biologically irrelevant.
To verify this expectation, we determined the
relative position of each SNP in a protein,
as shown in figure \ref{fig:snp_rel_pos}.

Per SNP, the 'protein' NCBI database was queried for the
protein sequence.
Of each protein sequence, the protein topology was determined 
using PureseqTM.
From the topology and the known location of the SNP, 
we score the location (i.e. TMH or soluble protein region) 
where the change occurred.

@@

@@

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Minor methods}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

PureseqTM does not predict the topology
of proteins that have less than three amino acids. 
The TRDD1 ('T cell receptor delta diversity 1') protein,
however, is two peptides long. 
The R package \verb;pureseqtmr;, however, 
predicts that mono- and di-peptides are cytosolic.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Minor discussion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \paragraph{Bacteria have a different cell membrane}

In this experiment we predicted epitopes that overlap with 
TMHs from a human, bacterial and viral proteome,
would these proteins be expressed in a human host.
Bacteria, however have different cell membranes and cell walls, 
hence different structural requirements for a TMH.
Both topology prediction tools were trained to recognize
human TMHs, thus we cannot be sure that
the transmembrane regions predicted in bacterial proteins
are actually part of an integral membrane protein.
For the purpose of this study, we assume the 
error in topology predictions to be unbiased way towards topology.
In other words: that a bacterial TMH is incorrectly
predicted to be absent just as often as it is incorrectly
predicted to be present elsewhere.

% \paragraph{False positives in SNPs}

Regarding the evolutionary conservation of TMHs using SNPs,
again, it is estimated that approximately ten percent
of SNPs is a false positive that result from the methods to determine
a SNP. One example is that sequence variations are incorrectly
detected due to highly similar duplicated sequences \cite{musumeci2010single}.
We assume that these duplications occur as often in TMHs as in
regions around these, hence we expect this not to affect our results.

@@

% \paragraph{Relative position of SNPs}

To verify if SNPs were sampled uniformly
over proteins, we show the distribution 
of the relative position in figure \ref{fig:snp_rel_pos}.
We find no clear evidence of a bias.
